Optimal (Multiway) Spatial Joins

### 2024-05-23 msj.tex
yufei

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$. The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. This problem holds significant importance in spatial databases and has been extensively studied for over two decades. We explain how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
\end{abstract}

@@

The proof is presented in Appendix~\ref{app:lmm:hv:type1:proj}. Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$. Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time.

@@

@@

\mytab{
    \> {\bf algorithm generate-$S$} \\
    \> 1.\> $S = \emptyset$ \\ 
    \> 2.\> {\bf for} each $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$ {\bf do} \\
    \> 3.\>\> define $B, r_a$, and $r_b$ as above \\
    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times \cross(r_b, B; V)$ {\bf do} \\
    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $S$
}

\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired. If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is
\myeqn{
    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) \nn
}
which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}. In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time. A symmetric algorithm can be used to compute all the $\cross(r_b, B; V)$ in the same complexity. This will then allow us to conclude that Step 2 can be done in $O(n \log n + |S|)$ time.

@@

\section{Proof of Lemma~\ref{lmm:hv:type1:proj}} \label{app:lmm:hv:type1:proj}

>>>>

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$. The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. This problem holds significant importance in spatial databases and has been extensively studied in the database community. We explain how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
\end{abstract}

@@

The proof can be found in Appendix~\ref{app:hv}. Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$. Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time.

@@

Appendix~\ref{app:hv} proves the next property:

\begin{lemma} \label{lmm:hv:type1:cross-non-empty}
    Neither $\cross(r_a, B; H)$ nor $\cross(r_b, B; H)$ can be empty.
\end{lemma}

@@

\mytab{
    \> {\bf Algorithm Generate-$S$} \\
    \> 1.\> $S = \emptyset$ \\ 
    \> 2.\> {\bf for} each $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$ {\bf do} \\
    \> 3.\>\> define $B, r_a$, and $r_b$ as above \\
    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times \cross(r_b, B; V)$ {\bf do} \\
    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $S$
}

\noindent In Appendix~\ref{app:hv}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired. If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is
\myeqn{
    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) \nn
}
which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}. In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time. A symmetric algorithm can be used to compute all the $\cross(r_b, B; V)$ in the same complexity. This will then allow us to conclude that Step 2 can be done in $O(n \log n + |S|)$ time.

@@

\section{Completing the Proof of Lemma~\ref{lmm:hv}} \label{app:hv}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:proj}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-non-empty}.}

\extraspacing {\bf Correctness of Algorithm Generate-$S$.}
